### 2021-07-20 doc/thesis/chapters/dsp.tex
impl details
new diagram (matlab transmitter)
matlab transmitter
transmitrer
fix all newlines

Many more factors could be listed, and elaborations over the various modulation types fill entire books.
However, with all these different facts, there are things that they all have in common.
On the transmitter side, all these signals are transmitted in a band-limited, high frequency signal.
The band limitation needs to be given at the transmitting antenna, to minimize interference to other bands in the frequency spectrum.
The frequency needs to be high, to radiate the signal by an antenna.

On the receiver side, once the signal reaches the antenna of a receiver, there are further similarities.
The antenna signal needs to be amplified and filtered at first.
Both steps depend on the antenna characteristics, but usually the received signal is of relatively low power, and the antenna receives a wider range of frequencies than required.
The filter therefor selects the frequency area that is of interest.
Another reason for the filter is to avoid spectral replicas in the subsequent modulation, or down-conversion.
Here, the front-end often down-converts to a low intermediate frequency (IF), instead of directly going to baseband.
The intermediate frequency is then shifted down to baseband by a modulator later in the digital domain, which brings advantages in the downstream processing chain. %TODO check this or explain more which advantages (HF to BB directly --> i/q complex signal? ; HF to IF to BB --> real signal ) ??

These three steps - amplification, filtering and down-conversion to an IF - are often done in the analog domain, since digital circuits cannot work with such high frequencies.
After the analog signal is down-converted to baseband, it needs to be digitized.
This is done by an analog-digital-converter (ADC).

After the signal is processed to digital baseband, the received signal can finally be decoded to get the actual data content.
This decoding process depends on the respective modulation type that is used.\\

@@

The IF is usually set to a fixed frequency, which results in the advantage, that the subsequent IF filter can be optimized, since it does not have to be adapted in frequency.
Therefore, a superhet receiver is able to select narrow band signals, even if the bands' surrounding frequencies are being used by other signals.
This is one of the superhets' main advantages.
A disadvantage is that the signal noise around the image frequency band is translated into the IF output band.
Thus, the filter requirements are relatively high, since they need to deliver a strong attenuation in order to suppress noise outside the desired band.

@@

A direct-conversion receiver, as its name suggests, converts an RF signal to baseband directly.
It can also be called homodyne, synchrodyne, or zero-IF receiver.

In order to achieve direct down-conversion, the local oscillator of the mixer needs to be set exactly to the center frequency of the desired signal.

The main advantage of direct-conversion receivers is, that their output is an image-free signal.
This is due to the properties of the mixer.
The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
These effects are not explained into more detail here, since this is not the focus of this thesis.

@@

The process of down-conversion is a technique of amplitude modulation (AM).
It is used to convert an RF signal to a signal of lower frequency, while preserving its information content.
This is the explanation in the time domain.
However, it may be easier look at the process in the frequency domain.
In the frequency domain, the band-limited RF signal is shifted down to a lower frequency.
This lower frequency can either be zero, i.e. baseband, or an intermediate frequency, such as in the application of a superheterodyne receiver.

In order to achieve this, the incoming signal needs to be multiplicated with a carrier signal, that is generated locally in the receiver, by a local oscillator (LO).
Depending on the frequency $f_{LO}$ of this local carrier, the resulting signal will be shifted to baseband, or an IF.
Note, that in the case of down-conversion, the local oscillator frequency $f_{LO}$ is interchangeably denoted as carrier frequency $f_c$.

@@

The procedure that is used for down-conversion is an amplitude modulation (AM).
The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
Their respective amplitudes are $A_m$ and $A_c$.
This equation is depicted as a block diagram and signal spectrum in Fig.\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.

The signal $m(t)$ is a baseband signal.
It may be antenna signal that was filtered by a bandpass filter.
The received signal is a real signal, and thus its negative frequency spectrum is an exact mirror of the positive frequencies.
This fact can be explained in two simple equations.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sampling Theorem}

Signal processing is often done in the digital domain, using digital signals.
A digital signal is obtained by sampling an analog signal in a periodic interval, the so-called sampling frequency $f_s$.

The sampling frequency needs to meet the sampling theorem, in order to correctly represent the analog signal.
The sampling theorem, by Nyquist and Shannon mainly consists of the following formula.
\begin{equation}
  f_{s} > 2 \cdot f_{signal}
\end{equation}

In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
Any digital signal processing technique must hold the specification of the sampling theorem \cite[chpt. 4.2]{AlessioSilviaMaria2016DSPa}.

In case the theorem is violated, the effect of aliasing occurs on the generated signal.
The sampling process introduces spectral replicas at integer multiples of the sampling frequency $f_s$.
These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
Fig.\ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.

@@

Frequency modulation (FM) is a widely used standard to transmit data streams.
The probably best known usecase therefor is commercial broadcast radio, where an audio stream is transmitted.
Devices to receive these streams are available for low prices to the public.

@@

The instantanious frequency $f_i$ of an FM signal can be described as
\begin{equation}
  f_i = f_c + \Delta f \cdot m(t)
\end{equation}
where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
Simply looking at this equation, the frequency deviation, which is sometimes also called the swing, varies in the range between $f_c \pm \Delta f$.

Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
\begin{equation}
  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big)
  \label{equ:fm_func}
\end{equation}
where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
This sensitivity factor has a direct relationship with the modulation index $\beta$.
\begin{equation}
  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m}
\end{equation}
where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
% More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm

Formula \eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
A widely used example application of FM is broadcast radio, where audio streams are transmitted.
An audio stream can be described as a cosine wave.
Therefore, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
\begin{equation}
  m(t) = A_m\ cos(2 \pi f_m t).
\end{equation}
By inserting this message signal into the generic FM formula \eqref{equ:fm_func}, the final equation for an FM signal transforms into the following form.
\begin{equation}
  y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
\end{equation}
Several formulas, as well as their derivations are described according to \cite[pg.54-55]{MatinMohammadA2018CSfE}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Frequency Band}

The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108 MHz.
However, some countries only partially use the band \cite[RR5-39]{ItuRadioRegulations}.
The range is located within the so-called Very-High-Frequency (VHF) band, which is open to the public in terms of usage.
This means, that any transmitter may use the specified frequency range freely.
Austria allowed the legal usage in 2006 \cite{AustriaRundfunkgesetz2014}.
However, the transmission power needs to be limited, so that neighboring receivers are not disturbed in receiving any existing channels.
The European Commision specifies the power limit as 50 nW of effective radiated power (ERP) \cite{EuCommissionRadioSpectrumHarmonization}.
Another limitation is, that a transmitter must be capable to transmit on multiple center frequencies within the broadcasting band.
It must not be fixed to a single center frequency.
Through this specification, a transmitter is capable of switching its transmission center frequency, in order to avoid interference with another transmitters signal.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Frequency Spectrum of a Broadcast FM Channel}

@@

The mono audio stream is located between 30 Hz and 15 kHz.
This signal is built by the sum of the left and right audio channels.
%The lower limit at 30 Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
%This effect is one of the downsides of amplitude modulation (AM), where the carrier is located at the channels' center frequeny and requires a high transmission power. % TODO: check this fact
The upper limit at 15 kHz is chosen to maintain a sufficient spacing to the first subcarrier.

For audio streaming, as in FM broadcasting, this is not a limitation, since it is unlikely to have frequencies higher than 15 kHz in an audio stream.
Also, this already reaches the upper limit of the human ears' bandwith, so higher spectral parts will not be heard anyway.

@@

A signal that is constructed by the difference between the left and right audio channel is modulated on a 38 kHz subcarrier.
This subcarrier is an integer multiple of the pilot tone at 19 kHz, for practical reasons.
However, the 38 kHz carrier is suppressed and thus not visible in the received spectrum.
The modulation technique that is used for this spectral part is called dual-sideband suppressed-carrier (DSB-SC).
Even though the carrier is suppressed, it can still be recovered, because it is phase coherent with the 19 kHz pilot tone per definition.

@@

Considering all these parts in the spectrum, there is still free bandwidth available to use up to the maximum channel bandwitdh of 100 kHz in one sideband.
Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Emphasis and De-Emphasis}

TODO

@@

The radio frequency antenna signal usually needs to be amplified, bandpass-filtered, and down-converted to baseband, using subsampling and quadrature-mixing, or similar strategies.
This part of the signal processing chain is assumed to be working correctly and is out of the scope of this chapter. %TODO: write up/down conversion in other chapter?
The FM signal that is evaluated here is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q) signals are available.

%TODO: show downconversion path (maybe as a figure?)
%https://www.veron.nl/wp-content/uploads/2014/01/FmDemodulator.pdf

In the following sections, two digital FM demodulator variants are described.

@@

@@

\includepicture [0.75] {Frequency discriminator block diagram \cite{FigBdFreqDiscriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}

@@

The differentiated signal still consists of a FM part, as explained above, using Eq.\eqref{equ:fm_demod_discriminator}.
In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
A block diagram is shown in Fig.\ref{fig:bd_freq_discriminator}.

@@

%TODO: explain a PLL in some more detail (block diagram)
A Phase-Locked Loop (PLL) can also be used to transform an FM signal to an AM signal.
A PLL is a feedback loop that is usually used to generate an output signal that has a fixed phase reference to its input.
In the case of FM demodulation, the loop filter in the feedback branch is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{SchnyderHaller2002}

@@

This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.

So at this point the FM signal is converted to an AM signal.
Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.

@@

The central branch selects the pilot tone at 19 kHz with a bandpass filter.
This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the mono and difference signals' frequency spectra are allocated.
Afterwards, the pilot tone frequency is doubled to achieve a 38 kHz subcarrier.
This guarantees a phase-coherency between pilot and subcarrier, which is required to correctly shift the difference signal down to baseband.

>>>>

Many more factors could be listed, and elaborations over the various modulation types fill entire books.
However, with all these different facts, there are things that they all have in common.
On the transmitter side, all these signals are transmitted in a band-limited, high frequency signal.
The band limitation needs to be given at the transmitting antenna, to minimize interference to other bands in the frequency spectrum.
The frequency needs to be high, to radiate the signal by an antenna.
On the receiver side, once the signal reaches the antenna of a receiver, there are further similarities.
The antenna signal needs to be amplified and filtered at first.
Both steps depend on the antenna characteristics, but usually the received signal is of relatively low power, and the antenna receives a wider range of frequencies than required.
The filter therefor selects the frequency area that is of interest.
Another reason for the filter is to avoid spectral replicas in the subsequent modulation, or down-conversion.
Here, the front-end often down-converts to a low intermediate frequency (IF), instead of directly going to baseband.
The intermediate frequency is then shifted down to baseband by a modulator later in the digital domain, which brings advantages in the downstream processing chain. %TODO check this or explain more which advantages (HF to BB directly --> i/q complex signal? ; HF to IF to BB --> real signal ) ??
These three steps - amplification, filtering and down-conversion to an IF - are often done in the analog domain, since digital circuits cannot work with such high frequencies.
After the analog signal is down-converted to baseband, it needs to be digitized.
This is done by an analog-digital-converter (ADC).
After the signal is processed to digital baseband, the received signal can finally be decoded to get the actual data content.
This decoding process depends on the respective modulation type that is used.\\

@@

The IF is usually set to a fixed frequency, which results in the advantage, that the subsequent IF filter can be optimized, since it does not have to be adapted in frequency.
Therefore, a superhet receiver is able to select narrow band signals, even if the bands' surrounding frequencies are being used by other signals.
This is one of the superhets' main advantages.
A disadvantage is that the signal noise around the image frequency band is translated into the IF output band.
Thus, the filter requirements are relatively high, since they need to deliver a strong attenuation in order to suppress noise outside the desired band.\\

@@

A direct-conversion receiver, as its name suggests, converts an RF signal to baseband directly.
It can also be called homodyne, synchrodyne, or zero-IF receiver.\\

In order to achieve direct down-conversion, the local oscillator of the mixer needs to be set exactly to the center frequency of the desired signal.
The main advantage of direct-conversion receivers is, that their output is an image-free signal.
This is due to the properties of the mixer.
The strongest disadvantages are the LO leakage, which leads to a DC offset at the mixer output, and the flicker noise, also called \textit{1/f}-noise, which induces strong noise at low frequencies \cite{PassosFábio2020AHSo}.
These effects are not explained into more detail here, since this is not the focus of this thesis.

@@

The process of down-conversion is a technique of amplitude modulation (AM).
It is used to convert an RF signal to a signal of lower frequency, while preserving its information content.
This is the explanation in the time domain.
However, it may be easier look at the process in the frequency domain.
In the frequency domain, the band-limited RF signal is shifted down to a lower frequency.
This lower frequency can either be zero, i.e. baseband, or an intermediate frequency, such as in the application of a superheterodyne receiver.
In order to achieve this, the incoming signal needs to be multiplicated with a carrier signal, that is generated locally in the receiver, by a local oscillator (LO).
Depending on the frequency $f_{LO}$ of this local carrier, the resulting signal will be shifted to baseband, or an IF.
Note, that in the case of down-conversion, the local oscillator frequency $f_{LO}$ is interchangeably denoted as carrier frequency $f_c$.\\

@@

The procedure that is used for down-conversion is an amplitude modulation (AM).
The equation for a generic AM signal is $ S_{AM}(t) = A_m m(t) \cdot A_c c(t) $, where $m(t)$ represents the message signal, and $c(t)$ is the high-frequency carrier.
Their respective amplitudes are $A_m$ and $A_c$.
This equation is depicted as a block diagram and signal spectrum in Fig.\ref{fig:bd_down_conversion_spectrum}, where $S_{AM}$ is denoted as the baseband signal $m_{BB}$ of the original signal $m$.
The signal $m(t)$ is a baseband signal.
It may be antenna signal that was filtered by a bandpass filter.
The received signal is a real signal, and thus its negative frequency spectrum is an exact mirror of the positive frequencies.
This fact can be explained in two simple equations.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sampling Theorem}
\label{sec:SamplingTheorem}

Signal processing is often done in the digital domain, using digital signals.
A digital signal is obtained by sampling an analog signal in a periodic interval, the so-called sampling frequency $f_s$.
The sampling frequency needs to meet the sampling theorem, in order to correctly represent the analog signal.
The sampling theorem, by Nyquist and Shannon mainly consists of the following formula.
\begin{equation}
  f_{s} > 2 \cdot f_{signal}
\end{equation}

In words: The sampling frequency must be higher than twice the frequency of the signals' highest frequency content.
Any digital signal processing technique must hold the specification of the sampling theorem \cite[chpt. 4.2]{AlessioSilviaMaria2016DSPa}.
In case the theorem is violated, the effect of aliasing occurs on the generated signal.
The sampling process introduces spectral replicas at integer multiples of the sampling frequency $f_s$.
These replicas overlap, if $f_s$ does not meet the theorem, which results in signal distortion.
Fig.\ref{fig:aliasing_freq_spectrum} illustrates the effect of aliasing in the frequency spectrum \cite{ThyagarajanK.S2019ItDS}.

@@

Frequency modulation (FM) is a widely used standard to transmit data streams.
The probably best known usecase therefor is commercial broadcast radio, where an audio stream is transmitted.
Devices to receive these streams are available for low prices to the public.\\

@@

The instantanious frequency $f_i$ of an FM signal can be described as
\begin{equation}
  f_i = f_c + \Delta f \cdot m(t)
\end{equation}
where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
Simply looking at this equation, the frequency deviation, which is sometimes also called the swing, varies in the range between $f_c \pm \Delta f$.
Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
\begin{equation}
  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int m(t) dt \Big)
  \label{equ:fm_func}
\end{equation}
where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
This sensitivity factor has a direct relationship with the modulation index $\beta$.
\begin{equation}
  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m}
\end{equation}
where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
% More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
Formula \eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
A widely used example application of FM is broadcast radio, where audio streams are transmitted.
An audio stream can be described as a cosine wave.
Therefore, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
\begin{equation}
  m(t) = A_m\ cos(2 \pi f_m t).
\end{equation}
By inserting this message signal into the generic FM formula \eqref{equ:fm_func}, the final equation for an FM signal transforms into the following form.
\begin{equation}
  y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
\end{equation}
Several formulas, as well as their derivations are described according to \cite[pg.54-55]{MatinMohammadA2018CSfE}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Frequency Band}
\label{sec:FrequencyBand}

The frequency band that is used for FM broadcasting is defined worldwide and spans from 87.5 to 108 MHz.
However, some countries only partially use the band \cite[RR5-39]{ItuRadioRegulations}.
The range is located within the so-called Very-High-Frequency (VHF) band, which is open to the public in terms of usage.
This means, that any transmitter may use the specified frequency range freely.
Austria allowed the legal usage in 2006 \cite{AustriaRundfunkgesetz2014}.
However, the transmission power needs to be limited, so that neighboring receivers are not disturbed in receiving any existing channels.
The European Commision specifies the power limit as 50 nW of effective radiated power (ERP) \cite{EuCommissionRadioSpectrumHarmonization}.
Another limitation is, that a transmitter must be capable to transmit on multiple center frequencies within the broadcasting band.
It must not be fixed to a single center frequency.
Through this specification, a transmitter is capable of switching its transmission center frequency, in order to avoid interference with another transmitters signal.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Frequency Spectrum of a Broadcast FM Channel}
\label{sec:FrequencySpectrumOfABroadcastFmChannel}

@@

The mono audio stream is located between 30 Hz and 15 kHz.
This signal is built by the sum of the left and right audio channels.
%The lower limit at 30 Hz is to prevent the transmission of a direct current (DC) part, which would require a large amount of power in transmission.
%This effect is one of the downsides of amplitude modulation (AM), where the carrier is located at the channels' center frequeny and requires a high transmission power. % TODO: check this fact
The upper limit at 15 kHz is chosen to maintain a sufficient spacing to the first subcarrier.
For audio streaming, as in FM broadcasting, this is not a limitation, since it is unlikely to have frequencies higher than 15 kHz in an audio stream.
Also, this already reaches the upper limit of the human ears' bandwidth, so higher spectral parts will not be heard anyway.

@@

A signal that is constructed by the difference between the left and right audio channel is modulated on a 38 kHz subcarrier.
This subcarrier is an integer multiple of the pilot tone at 19 kHz, for practical reasons.
However, the 38 kHz carrier is suppressed and thus not visible in the received spectrum.
The modulation technique that is used for this spectral part is called dual-sideband suppressed-carrier (DSB-SC).
Even though the carrier is suppressed, it can still be recovered, because it is phase coherent with the 19 kHz pilot tone per definition.\\

@@

Considering all these parts in the spectrum, there is still free bandwidth available to use up to the maximum channel bandwitdh of 100 kHz in one sideband.
Because of that, additional services were added to the pure audio transmission, to provide additional data services and information.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Pre-Emphasis and De-Emphasis}
%\label{sec:PreEmphasisAndDeEmphasis}
%TODO

@@

The radio frequency antenna signal usually needs to be amplified, bandpass-filtered, and down-converted to baseband, using subsampling and quadrature-mixing, or similar strategies.
This part of the signal processing chain is assumed to be working correctly and is out of the scope of this chapter. %TODO: write up/down conversion in other chapter?
The FM signal that is evaluated here is assumed to be a quadrature-mixed signal in baseband, which means that inphase and quadrature (I/Q) signals are available.
%
%TODO: show downconversion path (maybe as a figure?)
%https://www.veron.nl/wp-content/uploads/2014/01/FmDemodulator.pdf
%
In the following sections, two digital FM demodulator variants are described.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Baseband Delay Demodulator}
\label{sec:BasebandDelayDemodulator}
TODO\\
see /home/mike/Dropbox/Studium/ma\_semester\_4\_SS21/ma/literature/FmDemodulator.pdf

@@

\includepicture [0.75] [0] [H] {Frequency discriminator block diagram \cite{FigBdFreqDiscriminator}.} {bd_freq_discriminator} {img/draw.io/bd_freq_discriminator}

@@

The differentiated signal still consists of a FM part, as explained above, using Eq.\eqref{equ:fm_demod_discriminator}.
In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
A block diagram is shown in Fig.\ref{fig:bd_freq_discriminator}.\\

@@

%TODO: explain a PLL in some more detail (block diagram)
A Phase-Locked Loop (PLL) can also be used to transform an FM signal to an AM signal.
A PLL is a feedback loop that is usually used to generate an output signal that has a fixed phase reference to its input.
In the case of FM demodulation, the loop filter in the feedback branch is configured to be able to follow the frequency variations on the input, which is directly related to the encoded information.
The output of the PLL directly delivers an AM signal, which corresponds to the transmitted information. \cite{SchnyderHaller2002}\\

@@

This section describes the demodulation of the information content of an FM broadcast channel, as it is used in commercial FM radio broadcasting.
The assumption here is, that the actual FM demodulation, as described in chapter \ref{sec:algorithms-for-digital-fm-demodulation} is already done correctly, including the de-emphasis of the signal.
So at this point the FM signal is converted to an AM signal.
Thus, traditional AM DSP techniques can be employed to decode the FM broadcast channel.\\

@@

The central branch selects the pilot tone at 19 kHz with a bandpass filter.
This bandpass needs to have a frequency response that is sharp enough to have a sufficient attenuation at $\pm$4 kHz, since this is where the mono and difference signals' frequency spectra are allocated.
Afterwards, the pilot tone frequency is doubled to achieve a 38 kHz subcarrier.
This guarantees a phase-coherency between pilot and subcarrier, which is required to correctly shift the difference signal down to baseband.\\

### 2021-07-20 doc/thesis/chapters/introduction.tex
v-model
direct impl
new diagrams
remove unused/old pictures in img/
fix all newlines

In the master's degree program Embedded Systems Design a wide range of topics relate to digital signal processing, chip design and software development.
All of these disciplines find their application in an embedded system.
Multiple different methods of implementation are introduced for each topic and their usage is tested in practical sessions.
However, in the setting of a university course it is not always possible to look at the currently handled topic as a part of a larger, integrated system.
Instead, it is often treated as a standalone part that fulfills its own, specific functionality.

Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods in a common context of a single project.\\

@@

The main objective of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics which are accumulated over the time of the university program, in a single project.

This includes the development of a system model in a tool like Matlab, the practical usage and implementation of DSP in an FPGA, the development of a digital receiver and the deployment on actual hardware.\\

The receiver implementation on the FPGA is to be done in two variants.
The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL, as it has been done for years.
For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level \cplusplus code.
These two methods are implemented and compared on the basis of usability and some metrics.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FEEDBACK, Meeting 07-16-2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%formatting:
%- absätze überall gleich (mit oder ohne leerzeile)
%references:
%- hauptsache überall gleich (Fig.2.3 oder Sec.7.1 oder Eq.3 oder Chpt.4)
%- literature refs at the beginning of a chapter/section
%  like "this chapter is based on [1][4][5]."
%  or at the end of the last sentence like "...the end of this sentence. [7]"
%
%Allgemeiner Teil / Theorie:
% zusätzliches kapitel:
%   system design ansätze
%   - from algo to VHDL (book: vom algorithmus zur hardware)
%   - matlab to VHDL
%   - systementwuf richtung test driven design
%   --> design space exploration
%
%
% ganze Arbeit aufhängen an:
%   main objective: fit all the aspects (algorithm, impl, testing) into
%                   a single project
%
%
%
%

>>>>

In the master's degree program Embedded Systems Design a wide range of topics relate to digital signal processing, chip design and software development.
All of these disciplines find their application in an embedded system.
Multiple different methods of implementation are introduced for each topic and their usage is tested in practical sessions.
However, in the setting of a university course it is not always possible to look at the currently handled topic as a part of a larger, integrated system.
Instead, it is often treated as a standalone part that fulfills its own, specific functionality.
Therefore, one of the main motivation points of this thesis is to combine several disciplines and multiple implementation methods into a single system design, and treat it in this common project context.\\

@@

The main objective of this thesis is to elaborate a system design for an embedded system, that combines all the knowledge of different disciplines and topics which are accumulated over the time of the university program, in a single project.
This includes the development of a system model in a tool like Matlab, the practical usage and implementation of DSP in an FPGA, the development of a digital receiver and the deployment on actual hardware.\\

The receiver implementation on the FPGA is to be done in two variants.
The first variant is traditional, manually written hardware description language (HDL) in the programming language VHDL, as it has been done for years.
For the second variant, high-level synthesis (HLS) is employed, to generate the low-level HDL from high-level \cplusplus\ code.
These two methods are implemented and compared on the basis of usability and some metrics.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FEEDBACK, Meeting 07-16-2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%formatting:
%- absätze überall gleich (mit oder ohne leerzeile)
%references:
%- hauptsache überall gleich (Fig.2.3 oder Sec.7.1 oder Eq.3 oder Chpt.4)
%- literature refs at the beginning of a chapter/section
%  like "this chapter is based on [1][4][5]."
%  or at the end of the last sentence like "...the end of this sentence. [7]"
%
%Allgemeiner Teil / Theorie:
% zusätzliches kapitel:
%   system design ansätze
%   - from algo to VHDL (book: vom algorithmus zur hardware)
%   - matlab to VHDL
%   - systementwuf richtung test driven design
%   --> design space exploration
%
%
% ganze Arbeit aufhängen an:
%   main objective: fit all the aspects (algorithm, impl, testing) into
%                   a single project
%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO General
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ## unify all graphs and diagrams
%       - all gray; use exact same gray tone; eliminate the yellow?
%
%

### 2021-07-20 doc/thesis/chapters/system-design-process.tex
fix all newlines

The acronym SIMILAR consists of the following components.

\begin{itemize}
  \item \textbf{S}tate the Problem
  \item \textbf{I}nvestigate Alternatives
  \item \textbf{M}odel the System
  \item \textbf{I}ntegrate Components
  \item \textbf{L}aunch the System
  \item \textbf{A}ssess Performance
  \item \textbf{R}e-evaluate
\end{itemize}

@@

Multiple implementation methods are introduced in above sections, which all have different levels of abstraction.
VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
In Fig.\ref{fig:fpga_design_levels}, the various implementation methods are compared, with their respective levels of abstraction.\\

>>>>

\noindent
The acronym SIMILAR consists of the following components.

\begin{itemize}
  \item \textbf{S}tate the Problem
  \item \textbf{I}nvestigate Alternatives
  \item \textbf{M}odel the System
  \item \textbf{I}ntegrate Components
  \item \textbf{L}aunch the System
  \item \textbf{A}ssess Performance
  \item \textbf{R}e-evaluate\\
\end{itemize}

@@

Multiple implementation methods are introduced in above sections, which all have different levels of abstraction.
VHDL and Verilog require, or allow, a very detailed description of the RTL code, while SystemC, \cplusplus, or Matlab and Simulink provide a high level of abstraction.
In Fig.\ref{fig:fpga_design_levels}, the various implementation methods are compared.
The respective level of abstraction is higher towards the upper end, while the level of detail increases towards the bottom of the figure.\\
